Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-13 paper.qmd
Updated the Five Gaussian cluster projection

## Compute hexbin parameters
num_bins_x_gau1 <- 23

@@

```{r}
#| label: five-gau-tsne-projs
projection <- cbind(
  c(0.00660,-0.34942,0.14548,-0.12289),
  c(-0.18758,0.15398,0.30233,-0.08999))

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 2, 
               highd_data = scaled_gau_data, 
               model_highd = scaled_gau_data_model, 
               trimesh_data = tr_from_to_df_gau1, 
               axis_param = list(limits = 1,
                                 axis_scaled = 2,
                                 axis_pos_x = -0.72, 
                                 axis_pos_y = -0.72, 
                                 threshold = 0.1))

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 75

@@

```{r}
#| label: five-gau-umap-projs
projection <- cbind(
  c(0.00660,-0.34942,0.14548,-0.12289),
  c(-0.18758,0.15398,0.30233,-0.08999))

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 2, 
               highd_data = scaled_gau_data, 
               model_highd = scaled_gau_data_model, 
               trimesh_data = tr_from_to_df_gau1, 
               axis_param = list(limits = 1,
                                 axis_scaled = 2,
                                 axis_pos_x = -0.72, 
                                 axis_pos_y = -0.72, 
                                 threshold = 0.1))

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 33

@@

```{r}
#| label: five-gau-pacmap-projs
## First projection
projection <- cbind(
  c(0.00660,-0.34942,0.14548,-0.12289),
  c(-0.18758,0.15398,0.30233,-0.08999))

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 2, 
               highd_data = scaled_gau_data, 
               model_highd = scaled_gau_data_model, 
               trimesh_data = tr_from_to_df_gau1, 
               axis_param = list(limits = 1,
                                 axis_scaled = 2,
                                 axis_pos_x = -0.72, 
                                 axis_pos_y = -0.72, 
                                 threshold = 0.1))

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 2)

>>>>

## Compute hexbin parameters
num_bins_x_gau1 <- 18

@@

```{r}
#| label: five-gau-tsne-projs
projection <- cbind(
  c(0.10394,0.08379,-0.17868,0.32960),
  c(0.35111,-0.01137,-0.09657,-0.16018))

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 2.5, 
               highd_data = scaled_gau_data, 
               model_highd = scaled_gau_data_model, 
               trimesh_data = tr_from_to_df_gau1, 
               axis_param = list(limits = 1,
                                 axis_scaled = 2,
                                 axis_pos_x = -0.72, 
                                 axis_pos_y = -0.72, 
                                 threshold = 0.1))

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 63

@@

```{r}
#| label: five-gau-umap-projs
projection <- cbind(
  c(0.10394,0.08379,-0.17868,0.32960),
  c(0.35111,-0.01137,-0.09657,-0.16018))

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 2.5, 
               highd_data = scaled_gau_data, 
               model_highd = scaled_gau_data_model, 
               trimesh_data = tr_from_to_df_gau1, 
               axis_param = list(limits = 1,
                                 axis_scaled = 2,
                                 axis_pos_x = -0.72, 
                                 axis_pos_y = -0.72, 
                                 threshold = 0.1))

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 28

@@

```{r}
#| label: five-gau-pacmap-projs
## First projection
projection <- cbind(
  c(0.10394,0.08379,-0.17868,0.32960),
  c(0.35111,-0.01137,-0.09657,-0.16018))

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 2.5, 
               highd_data = scaled_gau_data, 
               model_highd = scaled_gau_data_model, 
               trimesh_data = tr_from_to_df_gau1, 
               axis_param = list(limits = 1,
                                 axis_scaled = 2,
                                 axis_pos_x = -0.72, 
                                 axis_pos_y = -0.72, 
                                 threshold = 0.1))

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 2)
